A tool set for random number generation on GPUs in R(?)

### 2021-11-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
random seeds in streams GPU
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2946 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG. If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory. \pkg{gpuRandom} is able to create streams both on the host and on the GPU device. \fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior. Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item. 
<<createStreams_CPU>>=
# creating streams on CPU
library("gpuR")
library("gpuRandom")
myStreamsCpu <- createStreamsCpu(n=4, seed=12345)
t(myStreamsCpu)
@
\fct{createStreamsCpu} has two arguments:
\begin{itemize}
\itemsep0em 
  \item \code{n}: number of streams to create.
  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
\end{itemize}
\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device. We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}. Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
<<createStreams_Gpu>>=
# creating streams on GPU
myStreamsGpu1 = vclMatrix(myStreamsCpu)
myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)
range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
@
\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''. \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}. By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.

>>>>

The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG. If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory. \pkg{gpuRandom} is able to create streams both on the host and on the GPU device. \fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior. Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item. 
<<createStreams_CPU>>=
# creating streams on CPU
library("gpuR")
library("gpuRandom")
myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
t(myStreamsCpu)
@
\fct{createStreamsCpu} has two arguments:
\begin{itemize}
\itemsep0em 
  \item \code{n}: number of streams to create.
  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
\end{itemize}
\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device. We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}. Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
<<createStreams_Gpu>>=
# creating streams on GPU
myStreamsGpu1 = vclMatrix(myStreamsCpu)
myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
@
\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''. \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}. By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.
